Everything you wanted to know about PanDA but were truly truly afraid to ask

### 2017-03-10 overall.tex
Iteration V

\begin{itemize}
    \item Comparison hardware performance Titan/Grid
    \item problem with lustre addressed with ramdisk
    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers.
    \item Competition for cache in AMD processors.
\end{itemize}

Currently, two main parameters measure the performance of the detector
simulation jobs submitted to Titan: (i) the time taken to setup
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit; and (ii) the distribution of the time taken by the
Geant4 toolkit to simulate a certain number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is  assembled from a large number of shared libraries. Once
bootstrapped, every algorithm and service is configured by a set of Python
scripts. Both these operations result in a large number of read operations,
including those required to  access of small python scripts.

@@

Originally, all shared libraries of AthenaMP and the python scripts for job
configuration were stored on the Spider 2 Lustre file system. However, the I/O
patterns of the initialization and configuration stages degraded the performance
of the filesystem. Since Spider 2 is a center-wide file system, shared by all
OLCF resources and users, this resulted in lower overall interactive and
metadata performance at OLCF\@. \mtnote{I am afraid the details about the trace
are too specific given the space constraints of a SC submission. Please feel
free to uncomment it if you disagree.}
% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
% exhibits a spike corresponding to the beginning of the runs before tapering
% off.

@@

This issue was addressed by moving the AthenaMP distribution to a read-only NFS
directory made accessible from the Titan work nodes. This eliminated the
metadata contention problem improving metadata read performance of two orders of
magnitude: from \~6,300 seconds on Lustre to \~1,500 seconds on NFS. Further, at
configuration time, the input files of each job were stored to a ramdisk on the
work nodes. This offered three orders of magnitude performance improvement in
the file validation step: from 1,320 seconds on Lustre to 40 seconds on NFS.

@@

>>>>

\begin{itemize}
    \item problem with lustre addressed with ramdisk
    \item Competition for cache in AMD processors.
    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers.
    \item Comparison hardware performance Titan/Grid
\end{itemize}

We use two main parameters to measure the performance of the detector simulation
jobs submitted to Titan: (i) the time taken to setup
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit; and (ii) the distribution of the time taken by the
Geant4 toolkit to simulate a certain number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is assembled from a large number of shared libraries. Once initialized,
every algorithm and service of AthenaMP is configured by a set of Python
scripts. Both these operations result in a large number of read operations on
the filesystem, including those required to access of small python scripts.

@@

Initially, all shared libraries of AthenaMP and the configuration python scripts
were stored on the Spider 2 Lustre file system. However, the I/O patterns of the
initialization and configuration stages degraded the performance of the
filesystem (Figure~\ref{}). Since Spider 2 is a center-wide file system, this
resulted in lower overall interactive and metadata performance for all OLCF
resources and users. \mtnote{I am afraid the details about the trace are too
specific given the space constraints of a SC submission. Please feel free to
uncomment it if you disagree.}
% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
% exhibits a spike corresponding to the beginning of the runs before tapering
% off.

@@

This issue was addressed by moving the AthenaMP distribution to a read-only NFS
directory made accessible from the Titan work nodes. This eliminated the problem
of metadata contention, improving metadata read performance from \~6,300 seconds
on Lustre to \~1,500 seconds on NFS. Further, at configuration time, the input
files of each detector simulation job were stored in a ramdisk on the work
nodes. This offered a marked performance improvement: from 1,320 seconds on
Lustre to 40 seconds on NFS.

@@

At runtime, detector simulation jobs showed performance degradation when all the
16 cores of Titan's work nodes were used. As shows in Figure~\ref{}, the mean
runtime for 8 concurrent jobs computing 100 events is XXXs, with each job
executed in 8 distinct processes, one per core. The mean runtime for 16
concurrent jobs computing 100 events is instead XXXs. This can be explained by
looking at the architecture of the cache in the AMD Opteron 6274 CPUs. This
processor shares the level 1 instruction cache between couples of cores,
impacting the performance of \ldots computation.

The compute performance of Titan for detector simulation jobs is, on average,
lower that that of LCWG. Figure~\ref{} shows the difference in execution time of
the same type of simulation between Titan and \ldots. This is not an issue with
the design of PanDA Brokers and, more in general, with the migration of part of
the ATLAS workload to supercomputers: Titan has been active since 2012 and it is
to be expected that more modern CPUs offer better performance. A pilot project
is already in place to test PanDA Broker on Summit, the next generation
supercomputer that will become active in 2018.
